GVN-Hoist: Hoisting Computations from Branches

### 2016-08-08 hoist.tex
contrib
partition

@@

\section{Code hoisting}
Similar to most compiler optimization passes the code-hoisting pass can also be
divided into the following five steps:

@@

\subsection{Checking for legality}
Finding a hoisting point i.e., a common dominator of the set of instructions. If
such a dominator is found then checking if all the use-operands of the set of
instructions are available or not. In some cases if the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
passing the legality check. Checking that the side-effects of the computations
does not intersect with any side-effects between the instructions to be hoisted
and their hoisting point. For memory operations like loads, stores, calls etc.,
it is also required to check that all the paths from the hoisting point to the
end of the function should execute the exact instruction, in order to guarantee
correctness. Moreover, hoisting of memory operations should not be done on paths
which have indirect branch targets e.g., landing pad, case statements, goto
labels etc., because it becomes tricky to prove that all the paths from hoisting
point to the end of the function would execute the instruction.

@@

\subsubsection{Hoisting loads}
At the hoisting point the availability of operands to the load (an address) is
checked. If it is not available we try to compute the addresss if that is
possible.  Along the path, from current position of the load instruction through
the hoisting point, we find if there are aliasing stores. In those cases we
discard the candidate.

@@

\subsubsection{Optimistic approach}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We see a lot of computations hoisted (see
Section~\ref{sec:experimental-results}. In order to do that, we collect the GVN
of all the instructions in the function. After that we iterate on the list of
instructions having identical GVN. After that we find the common dominator
dominating all the identical computations and try to hoist the instruction in
that basic block, if feasible. Many times it is not possible to hoist all the
instructions to one common dominator, due to legality/profitability
constraints. In those cases, the algorithmn would partion the list of
computations a partially hoist a set of computations to their common
dominator. For more details see Section~\ref{subsec:partition}.

>>>>

\subsection{Contributions of this paper}
\begin{itemize}
\item A systematic approach to writing a compiler optimization pass.
\item gvn hoisting
\item partition algorithm
\item cost model to reduce live-range and reduce spills
\end{itemize}

@@

\section{Code hoisting}
For the code hoisting we follow a very systematic approach to writing a compiler
optimization pass. As such this pass can be divided into the following five
steps:

@@

\subsection{Checking for legality}
Finding a hoisting point i.e., a common dominator of the set of instructions. If
such a dominator is found then checking if all the use-operands of the set of
instructions are available or not. In some cases if the operands are not
available, it is possible to reinstantiate (remateralize) the use-operands, thus
passing the legality check. Checking that the side-effects of the computations
does not intersect with any side-effects between the instructions to be hoisted
and their hoisting point. For memory operations like loads, stores, calls etc.,
it is also required to check that all the paths from the hoisting point to the
end of the function should execute the exact instruction, in order to guarantee
correctness. Moreover, hoisting of memory operations is tricky on paths which
have indirect branch targets e.g., landing pad, case statements, goto labels
etc., because it becomes difficult to prove that all the paths from hoisting
point to the end of the function would execute the instruction. In our current
implementation we discard hoisting on such paths.

In the optimistic approach (Section-\ref{subsec:optimistic}), it is possible
that a common hoisting point of all the instructions is either too far away, or
not possible. In these cases, it is still possible to `partially' hoist a subset
of instructions by splitting the set of candidates and finding a closer hoisting
point for each subset. For more details see Section-\ref{subsec:partition}.

@@

\subsubsection{Hoisting loads}
The availability of operands to the load (an address) is checked at the hoisting
point. If it is not available we try to rematerialize the addresss if that is
possible.  Along the path, from current position of the load instruction through
the hoisting point, we find if there are aliasing stores. In those cases we
discard the candidate.

@@

\subsubsection{Optimistic approach}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We see a lot of computations hoisted (see
Section~\ref{sec:experimental-results}. In order to do that, we collect the GVN
of all the instructions in the function. After that we iterate on the list of
instructions having identical GVN. After that we find the common dominator
dominating all the identical computations and try to hoist the instruction in
that basic block, if feasible. Many times it is not possible to hoist all the
instructions to one common dominator, due to legality/profitability
constraints. In those cases, the algorithmn would partion the list of
computations a partially hoist a set of computations to their common
dominator. For more details see Section-\ref{subsec:partition}.
